---
title: "spafdad"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{spafdad}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(spafdad)
```

# Introduction

The spafdad package is used for aggregating spatial data based on
given spatial resolutions. The package contains raster data of the
Netherlands. For specific resolutions, these raster data contains cells
which can be used for the aggregation.

If the users provides a dataset, containing measurements and spatial
coordinates using the Dutch 'Rijksdriehoekstelsel' (RD), these
measurements can be aggregated for each cell in one of the raster data
grids with selected resolution.

By using the package a overview of the spatial variability a a certain
measurement can be obtained without the need to provide the exact
location. 

For each cell distribution statistics can be calculated, such as the
mean, median, standard deviation, etc. Also the amount observations in
each cell can be calculated. 

Based on the amount of observations in each cell, the user can decide
what the minimal amount of observations is needed, to not disclose any
individual information.

# Installation

The package can be installed from GitHub.com by using the following command:

```{r eval = FALSE}
devtools::install_github("jspijker/spafdad")
```

# Usage

First, a datasets is needed with coordinates and measurements. The
coordinates should be in the Dutch 'Rijksdriehoekstelsel' (RD) system.
The measurements can be any type of numerical data, such as soil parameters. in
the package, a (somewhat artificial) dataset is provided with lutum
parameters. This dataset is used in this example.

```{r}
data("dclutum")
head(dclutum)
```

To bind the cell numers to the dataset, the function `grid_cell` can
be used. This function requires the dataset and the spatial resolution
of the grid. The function returns the dataset with an extra column
containing the cell numbers.

If the dataset contains columns indicating date, year, addresses, LIMS
number or other information, they will be kept in the dataset.

```{r}
x <- grid_cell(dclutum, 1000)
head(x)
```

In the above example we use grid with a resolution of 1000 meters. The
available resolutions are:

```{r}
get_resolutions()
```

The result of grid_cell can be used to aggregate the data. The
function `aggregate_data` can be used for this. The function requires
the dataset and the column name of the measurements.

```{r}
x_agg <- summary_cell(x, "lutum")
head(x_agg)
```

In the above example the mean, median, standard deviation, and
quantiles are calculated for each cell. At the moment, only the cell
id can be used for the aggregation. If other factors are present in
the dataset, they can't be used for the aggregation. For these
parameters the data should be aggregated before using the function or
multiple datasets can be created for each factor.

The amount of observations in each cell is calculated. This amount can be
used to filter out cells with a low amount of observations. Fort example,
each must contain at least 5 observations.

```{r}

x_disc <- x_agg |>
    dplyr::filter(n > 5)

head(x_disc)
```
